The globular cluster system of the Auriga simulations

### 2019-07-09 AurigaGCS.tex
Attempt to motivate each figure, and  explain their take away message for Sec. 4.1

In this work we use state of the art simulations that produce realistic spiral 
galaxies at redshift zero for which several global properties are consistent with 
the observations. The simulations yield realistic Milky Way analogues, and such
galaxies are known to host globular cluster systems. The question thus naturally 
arises whether the simulations (also) yield a population of star particles that 
is consistent with the Milky Way and/or Andromeda globular cluster systems. In 
particular, we investigate the Auriga suite of cosmological zoom simulations 
(\citealt{2017MNRAS.467..179G}, and described in Sec.~\ref{sec:auriga}) because
a study of its GC system could serve as an ultimate test of the star formation 
model. The plan of the paper is as follows. We summarise the relevant characteristics 
of the Auriga simulations in section~\ref{sec:auriga}, followed by a summary of 
the observations of the Milky Way (MW) and Andromeda globular cluster system (GCS)
in section~\ref{sec:observations} that we use to compare our simulations to in
section~\ref{sec:results}. We discuss our findings in section~\ref{sec:discussion}
to come to our conclusions in section~\ref{sec:conclusions}.

@@

\subsection{Milky Way}
\label{sec:milkyway}
\citet[][2010 edition; hereafter H96e10]{1996AJ....112.1487H} provides a
catalogue\footnote{See \url{https://www.physics.mcmaster.ca/Fac_Harris/mwgc.dat}}
of the Milky Way globular cluster system that contains properties of
157 GCs. The authors initially estimated the size of the MW GCS to be 180~$\pm$~10,
thus, their catalogue to be ${\sim}$85\% complete. However, an additional 59 GCs
are claimed to have been discovered by various authors. The total number of GCs
in the MW might add up to 216 with recent estimates now anticipating an additional 
thirty GCs yet to be discovered \citep[e.g.][and references therein]{2018ApJ...863L..38R}.
We still use data from the Harris catalogue, but caution that it may (only) be
53-72\% complete. Specifically, the relevant data fields that we use from H96e10
are the metallicity [Fe/H], the Galactic distance components $X$, $Y$, and $Z$ (in
kpc)\footnote{In a Sun-centered coordinate system: $X$ points toward Galactic
center, $Y$ in direction of Galactic rotation, and $Z$ toward the North Galactic
Pole. We calculate the galactocentric radius $R_{\text{GC}}=\sqrt{(X-R_\odot)^2
+ Y^2 + Z^2}$, assuming the solar radius $R_\odot=8$~kpc.}, and absolute
magnitude in the V-band $M_V$. We use the latter to calculate mass-estimates by
assuming $M_{V,\odot}=4.83$ and a mass to light ratio $M/L_V = 1.7$~M/L$_{\odot}$,
the mean for MW clusters \citep{2005ApJS..161..304M}.

@@

We supplement the catalogue with age-estimates from isochrone fits to stars
near the main-sequence turnoff in 55 GCs \citep[][hereafter V13]{2013ApJ...775..134V}.
The mean value of the age-estimates in this data set is $11.9 \pm 0.1$~Gyr and
the dispersion is $0.8$~Gyr. Furthermore, only one of the 55 GC age-estimates is
below $10$~Gyr.

@@

The M31 GCS has a mean age of $11.0 \pm 0.2$~Gyr with a dispersion of $2.2$~Gyr,
and $24$ GCs have age-estimates below $10$~Gyr with a minimum age of $4.8$~Gyr.
We show a histogram of the age-estimates of the $55$ MW GCs in V13 and $88$ GCs
in M31 for which age-estimates are available in CR16, see Figure~\ref{fig:MW-M31-age}.

@@

\subsection{Metallicity distribution}
\label{sec:results_FeH}
We investigate whether the star formation model implemented in Auriga produces
metallicity distributions consistent with the MW (M31) GC system, and whether
the subgrid generates sufficient total mass at metallicities typical for the
MW (M31) GCS.

\begin{figure*}
    \includegraphics[width=\columnwidth]{{Au4-10_FeH_cleaner-trim}.png}
    \includegraphics[width=\columnwidth]{{Au4-4_and_Au4_21_FeH-trim}.png}
    \caption{
        \emph{Left:} Metallicity distribution of Au4-10 (bottom panel). We show
        the GG candicates in green. We split the GC candidates into two
        subpopulations, those that have formed insitu (blue), and those that have
        been accreted (red). The dotted green line shows all star particles.
        The solid purple (magenta) line in the top panel shows the GC system
        of the MW (M31). \emph{Right:} Au4-4 (bottom), and Au4-21 (top).
        \label{fig:Au4-10and21_FeH}
    }
\end{figure*}
Figure~\ref{fig:Au4-10and21_FeH} shows the normalized metallicity distribution
of \mbox{Au4-10}\footnote{The nomenclature is `Au' for Auriga, followed by the
resolution level (4) and halo number (10, indicating which set of initial
conditions was used the run).}, \mbox{Au4-21}, and \mbox{Au4-4}, showing the GC
candidates in green. The accreted GC candidates are shown in red, and the insitu 
subpopulation in blue. The top panel of the left panel shows the MW (M31) GC 
system in purple (magenta). We overplot a double Gaussian for the MW GCS (the 
purple dashed lines), adopting literature values of the mean $\mu$ and standard
deviation $\sigma$ of the metal-rich and metal-poor populations \citep[][p. 38]{Harris2001}.

We find that the age cut, on average, lowers the mean metallicity from $0.0$ to $-0.5$.
Furthermore, the old accreted stars generally have lower metallicities than the
old insitu stars. The difference between the mean metallicity of the old insitu
and old accreted stars for the majority of the simulations is ${\sim}0.3$ (e.g.
\mbox{Au4-21}), and ${\sim}0.5$ dex for \mbox{Au4-10}, \mbox{Au4-16}, \mbox{Au4-17},
\mbox{Au4-18} and \mbox{Au4-22}. This trend is only reversed for \mbox{Au4-1}
and \mbox{Au4-4}, for which the old insitu population has a lower mean metallicity
instead. For \mbox{Au4-1} we find $\mu = -1.51 \, ({-0.74})$ for the old insitu (accreted),
but the former consists of only $1019$ particles ($1.3$\% of all GC candidates, and
with a total mass of 5e7 M\Sun) thus we believe our classification of insitu is
flawed for this halo due to misclassification of the primary halo in the merger
tree by subfind. For \mbox{Au4-4}, ${10.8}$\% of the GC candidates is classified
as insitu (compared to insitu fractions of ${40-80}$\% for other haloes). Overall
we find that the simulations produce (sub)populations of GC candidates that are
more metal-rich than the MW and M31 GC systems. Moreover, none of the simulations
has a population of GC candidates with a bimodal metallicity distribution
(the green curves).

@@

We show the mean metallicity and standard deviation of all thirty Auriga L4 haloes
in Figure~\ref{fig:FeH_mu_sigma}. The green crosses are to be compared to the
purple (magenta) cross, which shows the mean value of all MW (M31) GCs.
We test the null hypothesis that MW (M31) and the GC candidates (all, insitu, and
accreted) are drawn from the same underlying distribution by calculating the KS
test statistic (i.e. six KS tests per simulation). We reject $H_0$ at the 100.00~\%
confidence level for virtually all subsets of all simulations, both for MW and M31,
except when comparing the old accreted stars of \mbox{Au4-10} to the MW GCS. In this
case we still reject $H_0$ with, but with $\leq 98.68$~\% confidence. When comparing
to the M31 GCS, the `best' matches are the old accreted subpopulations of \mbox{Au4-13},
\mbox{Au4-15}, and \mbox{Au4-17} ($H_0$ rejected with $\leq$ 99.82~\% confidence, 99.95~\%,
and 98.32~\%). Here we find that the cumulative distributions are more or less
similar up to [Fe/H]~$-0.5$, above which the simulations yield higher number
counts which drives the KS test statistic over the critical values. This affirms
our finding that the simulations produce GC candidates that are more metal-rich
than the MW (M31) GCS. The main reasons that $H_0$ is rejected with such
high confidence are under-production of old stars with [Fe/H]~$\leq$~-1.5,
and over-production of stars with [Fe/H]$\geq$~-0.5.

In addition, we show the metal-rich (metal-poor) population of the MW GCS using
a solid (open) dot. The mean metallicities of the old insitu populations appear
roughly consistent with that of the metal-rich population of the MW, but the
simulations show larger dispersions. We are uncertain whether definitive consensus
is reached concerning uni- or bimodality in the [Fe/H] distribution of M31, but
CR16 argues that the data, after removal of younger objects due to improved age
classification, hints at three populations. However, we do not show these (unknown)
points for the M31 GCS.

\begin{figure}
    \includegraphics[width=\columnwidth]{{logMFeH_withRatios-trim}.png}
    \caption{
        Mass-weighted metallicity distribution of star particles in the Auriga
        simulations. We show the median value of all Auriga haloes for all
        stars (green dotted) and globular cluster candidates (i.e. stars with
        age $>$~10~Gyr; green solid). The latter sub set is further split up
        between stars that formed in-situ (blue solid), and those that were accreted
        (red solid). Shaded regions indicate the $1\sigma$ interval. The MW (M31)
        GCS is shown in purple (magenta). The middle (bottom) panel shows the
        ratio of the simulated mass to the mass in the MW (M31) GCS.
        \label{fig:FeH}
    }
\end{figure}
We now turn to the total mass in GC candidates produced by the Auriga simulations.
We show the median (coloured lines) for all thirty Auriga L4 haloes with the
$1\sigma$ interval around it (shaded regions, which shows the scatter between
runs that have different initial conditions, thus, have unique merger histories).
We chose to aggregate the data to indicate a general trend that we find when the
GC candidates are split up according to birth location, rather than selecting
typical examples of individual (simulated) galaxies. Figure~\ref{fig:FeH} shows 
the mass-weighted metallicity distribution of all Auriga L4 galaxies. Once again 
we notice that the peak metallicity shifts down from $0$ to $-0.5$ for old stars
(green solid) compared to all stars (green dotted), and we learn that the mass at 
the peak lowers by roughly one dex. The mass budget of the old stars is dominated
by the old insitu population (blue solid) below [Fe/H]~=~$-1$, and by the old accreted
stars (red solid) above this value. We show the MW (M31) GCS in purple (magenta)
and we show the ratio of the simulated to the observed profiles in the middle
(bottom) panel. This mass excess can be thought of the `mass budget' that the
Auriga GC candidates can `afford to lose' due to a combination of smaller than
unity bound cluster formation efficiencies combined with a Hubble time of dynamical
evolution, while still producing a GC system with sufficient mass to be consistent
with the MW (M31) GCS. The cluster formation efficiency would have to linearly
decrease with decreasing metallicity for Auriga GC candidates to produce a population
of GC candidates that is consistent with the MW. For the GC candidates in M31
we find a constant mass excess up to $-0.9$, above which the
simulations produce a higher mass excess with increasing metallicity. If dynamical
evolution is not expected to more efficiently disrupt GCs of higher metallicity,
then we would find that the efficiency to form bound star clusters would have to
decrease with increasing metallicity.

>>>>

In this work we use state of the art simulations that produce realistic spiral 
galaxies at redshift zero for which several global properties are consistent with 
the observations. The simulations yield realistic Milky Way analogues, and real
galaxies host globular cluster systems. The question thus naturally arises whether 
the star formation histories of the simulations (also) give rise to a GC system 
similar to the Milky Way and/or Andromeda globular cluster systems. In particular, 
we use the Auriga suite of cosmological zoom simulations 
(\citealt{2017MNRAS.467..179G}, and described in Sec.~\ref{sec:auriga}) to
investigate whether the star formation model implemented produces metallicity, 
radial, and metallicity-radial distributions that are consistent with the MW 
and/or M31 GC systems, and whether the model produces enough stellar mass with
the right properties to allow for formation efficiencies lower than unity and
the expected (dynamical) mass loss over the cluster lifetime (not modelled 
explicitly in this study).

The plan of the paper is as follows. We summarise the relevant characteristics 
of the Auriga simulations in section~\ref{sec:auriga}, followed by a summary of 
the observations of the Milky Way (MW) and Andromeda globular cluster system (GCS)
in section~\ref{sec:observations} that we use to compare our simulations to in
section~\ref{sec:results}. We discuss our findings in section~\ref{sec:discussion}
to come to our conclusions in section~\ref{sec:conclusions}.

@@

\subsection{Milky Way}
\label{sec:milkyway}
\citet[][2010 edition; hereafter H96e10]{1996AJ....112.1487H} provides a
catalogue\footnote{See \url{https://www.physics.mcmaster.ca/Fac_Harris/mwgc.dat}}
of the Milky Way globular cluster system that contains properties of
157 GCs. The authors initially estimated the size of the MW GCS to be 180~$\pm$~10,
thus, their catalogue to be ${\sim}$85\% complete. However, an additional 59 GCs
are claimed to have been discovered by various authors. The total number of GCs
in the MW might add up to 216 with recent estimates now anticipating an additional 
thirty GCs yet to be discovered \citep[e.g.][and references therein]{2018ApJ...863L..38R}.
We still use data from the Harris catalogue, but caution that it may (only) be
53-72\% complete. Specifically, the relevant data fields that we use from H96e10
are the metallicity [Fe/H], the Galactic distance components $X$, $Y$, and $Z$ (in
kpc)\footnote{In a Sun-centered coordinate system: $X$ points toward Galactic
center, $Y$ in direction of Galactic rotation, and $Z$ toward the North Galactic
Pole. We calculate the galactocentric radius $R_{\text{GC}}=\sqrt{(X-R_\odot)^2
+ Y^2 + Z^2}$, assuming the solar radius $R_\odot=8$~kpc.}, and absolute
magnitude in the V-band $M_V$. We use the latter to calculate mass-estimates by
assuming $M_{V,\odot}=4.83$ and a mass to light ratio $M/L_V = 1.7$~M/L$_{\odot}$,
the mean for MW clusters \citep{2005ApJS..161..304M}. We supplement the catalogue
with age-estimates from isochrone fits to stars near the main-sequence turnoff 
in 55 GCs \citep[][hereafter V13]{2013ApJ...775..134V}.

@@

@@

\subsection{Age-estimates}
Figure~\ref{fig:MW-M31-age} shows a histogram of the age-estimates of the $55$
MW GCs in V13 and $88$ GCs in M31 for which age-estimates are available in CR16.
The mean age of the MW GCs in this data set is $11.9 \pm 0.1$~Gyr and the 
dispersion is $0.8$~Gyr. Furthermore, only one of the 55 GC age-estimates is
below $10$~Gyr. The M31 GCS has a mean age of $11.0 \pm 0.2$~Gyr with a dispersion 
of $2.2$~Gyr, and $24$ GCs have age-estimates below $10$~Gyr with a minimum age 
of $4.8$~Gyr. The Milky Way appears to host a globular cluster system that is
somewhat older than the Andromeda galaxy, although we caution that both data 
sets are incomplete and the age measurements have large uncertainties (of $1-2$~Gyr)

@@

\subsection{Metallicity distribution}
\label{sec:results_FeH}
We investigate whether the star formation model implemented in Auriga produces
metallicity distributions consistent with the MW and M31 GC systems, and whether
the subgrid generates sufficient total mass at metallicities typical for the
MW and M31 GCS's. To visually inspect the former we show the normalized metallicity 
distribution of three specific Auriga galaxies in Figure~\ref{fig:Au4-10and21_FeH}
in comparison to the MW and M31 [Fe/H] distributions.

\begin{figure*}
    \includegraphics[width=\columnwidth]{{Au4-10_FeH_cleaner-trim}.png}
    \includegraphics[width=\columnwidth]{{Au4-4_and_Au4_21_FeH-trim}.png}
    \caption{
        \emph{Left:} Metallicity distribution of Au4-10 (bottom panel). We show
        the GG candicates in green. We split the GC candidates into two
        subpopulations, those that have formed insitu (blue), and those that have
        been accreted (red). The dotted green line shows all star particles.
        The solid purple (magenta) line in the top panel shows the GC system
        of the MW (M31). \emph{Right:} Au4-4 (bottom), and Au4-21 (top).
        \label{fig:Au4-10and21_FeH}
    }
\end{figure*}
We select \mbox{Au4-10}\footnote{The nomenclature is `Au' for Auriga, followed by the
resolution level (4) and halo number (10, indicating which set of initial
conditions was used the run).}, \mbox{Au4-21}, and \mbox{Au4-4} to show the
distributions of three individual simulation runs that we consider representative
for specific behaviour, and to highlight that different runs could give rise to 
different distributions. We plot the age-selected GC candidates in green, those
that are accreted in red, and the insitu subpopulation in blue. The top 
half of the left figure shows the MW (M31) GC system in purple (magenta), where 
we overplot a double Gaussian for the MW GCS (the purple dashed lines) by 
adopting literature values of the mean $\mu$ and standard deviation $\sigma$ of 
the metal-rich and metal-poor subpopulations \citep[][p. 38]{Harris2001}.

We find that the age cut, on average, lowers the mean metallicity from $0.0$ to
$-0.6$. Furthermore, the old accreted stars generally have lower mean metallicities
than the old insitu stars with differences of roughly ${\sim}0.3$ for the majority 
of the simulation runs. This behaviour can be seen in \mbox{Au4-21}, while a 
slightly larger difference of ${\sim}0.5$ dex is seen for \mbox{Au4-10} (as plotted), 
\mbox{Au4-16}, \mbox{Au4-17}, \mbox{Au4-18} and \mbox{Au4-22}. However, this trend 
is reversed for \mbox{Au4-1} and \mbox{Au4-4} for which the old insitu population 
has a lower mean metallicity instead. We caution that \mbox{Au4-1} is undergoing
a major merger at redshift zero and note that we find $\mu = -1.51 \, ({-0.74})$ 
for the old insitu (accreted) although the former consists of only $1019$ 
particles ($1.3$\% of all GC candidates, and with a total mass of 5e7 M\Sun). 
For \mbox{Au4-4}, ${10.8}$\% of the GC candidates is classified as insitu (compared 
to insitu fractions of ${40-80}$\% for other haloes). Overall we find that the 
simulations produce (sub)populations of GC candidates that are
more metal-rich than the MW and M31 GC systems. Moreover, none of the simulations
has a population of GC candidates with a bimodal metallicity distribution
(the green curves).

@@

We show the mean metallicity and standard deviation of all thirty Auriga L4 haloes
in Figure~\ref{fig:FeH_mu_sigma} to show the behaviour for the full set of Auriga
galaxies. The green crosses are to be compared to the purple (magenta) cross, 
which shows the mean value of all MW (M31) GCs. In addition, we show the metal-rich 
(metal-poor) population of the MW GCS using a solid (open) dot. We caution that
these literature values result from individual Gaussian fits to sub sets of the 
observational data cut at [Fe/H]~$= -1$. We do not include corresponding data 
points calculated using such an artificial cut for the simulated star particles.
Having said that, we do find it interesting that the mean metallicities of the 
old insitu populations appear roughly consistent with that of the metal-rich 
population of the MW although the simulations show larger dispersions. The latter 
could simply be caused by the hard separation of the data into two groups, which 
means the range is smaller and the resulting dispersion could be lower. With 
regard to the M31 GCS, we are uncertain whether definitive consensus is reached
concerning uni- bi- or trimodality in the [Fe/H] distribution, but CR16 argues 
that the data, after removal of younger objects due to improved age classification, 
hints at three populations separated at [Fe/H]~$>-0.4$ and [Fe/H]~$ < 1.5$. None 
of the Auriga sub sets has a mean value that offers much hope that the simulated 
distribution is consistent with the lowest metallicity group in the M31 GCS. This 
is also true for the metal-poor population of the MW with a mean value of $-1.6$.
The main take-away from this plot, again, is that all Auriga L4 galaxies have
metallicity distributions with (much) larger mean values than what is observed
for the MW and M31 GCS, and that we systematiclly find lower mean metallicities
for the accreted GC candidates than for those that have formed in situ.

Finally, we test the null hypothesis that MW (M31) and the GC candidates (all, insitu, and
accreted) are drawn from the same underlying distribution by calculating the KS
test statistic (i.e. six KS tests per simulation). We reject $H_0$ at the 100.00~\%
confidence level for virtually all subsets of all simulations, both for MW and M31,
except when comparing the old accreted stars of \mbox{Au4-10} to the MW GCS. In this
case we still reject $H_0$ with, but with $\leq 98.68$~\% confidence. When comparing
to the M31 GCS, the `best' matches are the old accreted subpopulations of \mbox{Au4-13},
\mbox{Au4-15}, and \mbox{Au4-17} ($H_0$ rejected with $\leq$ 99.82~\% confidence, 99.95~\%,
and 98.32~\%). Here we find that the cumulative distributions are more or less
similar up to [Fe/H]~$-0.5$, above which the simulations yield higher number
counts which drives the KS test statistic over the critical values. This affirms
our finding that the simulations produce GC candidates that are more metal-rich
than the MW (M31) GCS. The main reasons that $H_0$ is rejected with such
high confidence are under-production of old stars with [Fe/H]~$\leq$~-1.5,
and over-production of stars with [Fe/H]$\geq$~-0.5.

\begin{figure}
    \includegraphics[width=\columnwidth]{{logMFeH_withRatios-trim}.png}
    \caption{
        Mass-weighted metallicity distribution of star particles in the Auriga
        simulations. We show the median value of all Auriga haloes for all
        stars (green dotted) and globular cluster candidates (i.e. stars with
        age $>$~10~Gyr; green solid). The latter sub set is further split up
        between stars that formed in-situ (blue solid), and those that were accreted
        (red solid). Shaded regions indicate the $1\sigma$ interval. The MW (M31)
        GCS is shown in purple (magenta). The middle (bottom) panel shows the
        ratio of the simulated mass to the mass in the MW (M31) GCS.
        \label{fig:FeH}
    }
\end{figure}
We now turn to the second goal, which is to see whether sufficient total mass in 
GC candidates produced by the Auriga simulations. To answer this question we plot
a mass-weighted metallicity distribution in Figure~\ref{fig:FeH}. We show the 
median (coloured lines) for all thirty Auriga L4 haloes with the $1\sigma$ 
interval around it (shaded regions, which shows the scatter between runs that 
have different initial conditions, thus, have unique merger histories). We chose 
to aggregate the data to indicate general trends that we find when the GC candidates 
are split up according to birth location, rather than selecting typical examples 
of individual (simulated) galaxies. Once again we notice that the peak metallicity 
shifts down from $0$ to $-0.6$ for old stars (green solid) compared to all stars 
(green dotted), and we learn that the mass at the peak lowers by roughly one dex. 
The mass budget of the old stars is dominated by the old insitu population (blue 
solid) below [Fe/H]~=~$-1$, and by the old accreted stars (red solid) above this 
value. We show the MW (M31) GCS in purple (magenta) and we show the ratio of the 
simulated to the observed profiles in the middle (bottom) panel. This mass excess 
can be thought of the `mass budget' that the Auriga GC candidates can `afford to 
lose' due to a combination of smaller than unity bound cluster formation 
efficiencies combined with a Hubble time of dynamical evolution, while still 
producing a GC system with sufficient mass to be consistent with the MW (M31) GCS. 
The cluster formation efficiency would have to linearly decrease with decreasing 
metallicity for Auriga GC candidates to produce a population of GC candidates 
that is consistent with the MW. For the GC candidates in M31 we find a constant 
mass excess up to $-0.9$, above which the simulations produce a higher mass excess 
with increasing metallicity. If dynamical evolution is not expected to more 
efficiently disrupt GCs of higher metallicity, then we would find that the 
efficiency to form bound star clusters would have to decrease with increasing 
metallicity.
